---
title: "SS_DEmodels"
author: "Laura Gramolini"
date: '2022-09-21'
output: html_document
---

Here we test the models
First we load the data putting the Host as design. I would load the data everytime and change the design, probably not the best but the easiest way.

```{r}
library(DESeq2)
library(tidyverse)
SScountdf <- read.csv("~/GitHub/Worms/Schistocephalus/data/SScountdf.csv", header = TRUE, sep="")
SScolData <- read.csv("~/GitHub/Worms/Schistocephalus/data/schistoData.txt", header = TRUE, sep = "")
```


M1 would be the dataset loaded in DEseq with the Host as design
```{r}
M1 <- DESeqDataSetFromMatrix(
  countData = SScountdf,
  colData = SScolData,
  design= ~ Host)
```

The we use DEseq with the likelihood ratio test to compare the Host model with the intercept

```{r}
HostM1 <- DESeq(M1, test="LRT", reduced = ~1)
```
Calculate the results
```{r}
resultsM1 <- results(HostM1)
```

Load the data with Stage as design: M2
```{r}
M2 <- DESeqDataSetFromMatrix(
  countData = SScountdf,
  colData = SScolData,
  design= ~ Stage)
```

Test Stage against intercept
```{r}
StageM2 <- DESeq(M2, test="LRT", reduced = ~1)
```

Calculate the results 
```{r}
resultsM2 <- results(StageM2)
```


Load the data with Host and Stage interaction (Condition) against the intercept: M3
```{r}
M3 <- DESeqDataSetFromMatrix(
  countData = SScountdf,
  colData = SScolData,
  design= ~ Condition)
```

Now Test the M3 model (Host * Stage) against the intercept
```{r}
ConditionM3 <- DESeq(M3, test="LRT", reduced = ~ 1)
resultsM3 <- results(ConditionM3)
```

Load the data with Design condition usign a matrix to then test against the Host and Stage
```{r}
condmatrix <- model.matrix(~Condition, SScolData)
M4 <- DESeqDataSetFromMatrix(
  countData = SScountdf,
  colData = SScolData,
  design= condmatrix)
```


Since we used a matrix for full model we have to use the matrix for Host
```{r}
hostmatrix <- model.matrix(~Host, SScolData)
```

Now we test the the condition and the Host
```{r}
ConditionM4 <- DESeq(M4, test = "LRT", reduced = hostmatrix)
```

```{r}
resultsM4 <- results(ConditionM4)
```

Now with the same data loaded in M4 (condmatrix) we test against the stage, so we need a stage matrix
```{r}
Stagematrix <- model.matrix(~Stage, SScolData)
```

And then we test
```{r}
ConditionM5 <- DESeq(M4, test = "LRT", reduced = Stagematrix)
resultsM5 <- results(ConditionM5)
```



Now I have to create a data frame to plot and compare the models. So I have to take the adjpvalues. I'll make 1 column with the genes, one column with the models and one column with the adjusted p values.

```{r}
int_host <- as.data.frame(resultsM1)
int_stage <- as.data.frame(resultsM2)
host_condition <- as.data.frame(resultsM4)
stage_condition <- as.data.frame(resultsM5)
int_condition <- as.data.frame(resultsM3)


DEGsmodels <- dplyr::lst(int_host, int_stage, host_condition, stage_condition, int_condition) %>% 
  set_names(c("int_host", "int_stage", "host_condition", "stage_condition", "int_condition")) %>% 
  bind_rows(.id = "Model") %>% 
  select(Model, padj)

write.table(DEGsmodels, file="~/GitHub/Worms/Schistocephalus/data/DEGsmodels.csv")
```


```{r}
filteredDEGs <- filter(DEGsmodels, padj < 0.00000000000001)
```


Plot the adjpvalues
```{r}
ggplot(DEGsmodels, aes(x=padj+0.0000000001, color = Model)) + 
         geom_density() +
  scale_x_log10()
```

```{r}
filter(DEGsmodels) %>%
ggplot(aes(x=padj+0.00000001, color = Model, fill = Model)) + 
         geom_histogram()+
  scale_x_log10()
```



```{r}
plotMA(resultsM1, ylim=c(-2,2))
```

```{r}
plotMA(resultsM3, ylim=c(-2,2))
```

```{r}
as.data.frame(resultsM3) %>% 
ggplot(aes(x=log2FoldChange, y=-log10(pvalue))) + 
         geom_point()
```

```{r}
as.data.frame(resultsM1) %>% 
ggplot() +
  geom_smooth(aes(x=pvalue, y=log2FoldChange))
```



```{r}
use <- resultsM1$baseMean > metadata(resultsM1)$filterThreshold
h1 <- hist(resultsM1$pvalue[!use], breaks=0:50/50, plot=FALSE)
h2 <- hist(resultsM2$pvalue[use], breaks=0:50/50, plot=FALSE)
colori <- c(`do not pass`="khaki", `pass`="powderblue")
```

```{r}
barplot(height = rbind(h1$counts, h2$counts), beside = FALSE,
        col = colori, space = 0, main = "", ylab="frequency")
text(x = c(0, length(h1$counts)), y = 0, label = paste(c(0,1)),
     adj = c(0.5,1.7), xpd=NA)
legend("topright", fill=rev(colori), legend=rev(names(colori)))
```


