---
title: "SS_QA"
author: "Laura Gramolini"
date: '2022-09-21'
output: html_document
---

Here we perform some quality check on the count matrix, following the DEseq pipeline

First we load the count matrix (SScountdf.csv)
```{r}
SScountdf <- read.csv("~/GitHub/Worms/data/SScountdf.csv", header = TRUE, sep = "")
```


```{r}
ggplot(SScountdf) +
  geom_histogram(aes(x = ConditionA03), stat = "bin", bins = 200) +
  xlab("Raw expression counts") +
  ylab("Number of genes")+
  scale_x_log10()

```
We create another table with genes, sample and count as columns (pivot)



Here we can build a graph with several samples for the same condition
```{r}
ggplot(pivot%>%
         filter(grepl("ConditionL", name)))+
  geom_histogram(aes(x = value), stat = "bin", bins = 200) +
  xlab("Raw expression counts") +
  ylab("Number of genes")+
  scale_x_log10()+
  facet_wrap(~name)

```

Load the metadata for Schistocephalus

```{r}
SScolData <- read.table("~/data/schistoData.txt")
```

And Now I can start loading the data in DEseq2 but first let's check:
```{r}
head(SchistoFCounts$counts)
```

Loading the tables in DEseq2

```{r}
str(SScolData)
SScolData$Host <- as.factor(SScolData$Host)
SScolData$Unique_name <- as.factor(SScolData$Unique_name)
SScolData$Stage <- as.factor(SScolData$Stage)
SScolData$Condition <- as.factor(SScolData$Condition)

SSDEseq <- DESeqDataSetFromMatrix(
  countData = SchistoFCounts$counts,
  colData = SScolData,
  design = ~1)
```

I've put no design here, cause I just want to check the data in the following steps
And no ref level


Let's remove the genes that are counted less than 10 times in all the samples
For now ok, but then decide the threshold base on a reason, why 10? why 20? For the sum of the samples? At least how many? Make several try. Always do things that you can justify
```{r}
keep <- rowSums(counts(SSDEseq)) >= 10
SSDEseq <- SSDEseq[keep,]
```


Quality checks on the data loaded

Let's plot sd vS mean of each gene to see if the range of variance is the same at different ranges of mean, so to see if they are homoskedastic:
```{r}
lambda <- 10^seq(from = -1, to = 2, length = 1000) ###here, Poisson counts with a range of lambda from 0.1 to 100
SSmatrix <- matrix(rpois(1000*100, lambda), ncol = 100)
library("vsn")
meanSdPlot(SSmatrix, ranks = TRUE)
```

They are not, and this is normal for RNA-seq data, cause lowly expressed genes will have a low range of variance, and this will reflect on the PCA or other exploratory analysis, so we need to transform the data:
I'm gonna apply the vst transformation cause I have more tha 50 samples and because is faster...
And I will run it blind so the variance doesn't get overestimated by the sample size
(look into it)
```{r}
SSDEseqvst <- vst(SSDEseq, blind = TRUE)
head(assay(SSDEseqvst), 3)
```

Let's see how it looks now

```{r}
meanSdPlot(assay(SSDEseqvst), rank=TRUE)
```

And plotting 2 random genes:
```{r}
SSvstdf <- as_data_frame(assay(SSDEseqvst)[, 1:2]) %>% mutate(transformation = "vst")
colnames(SSvstdf)[1:2] <- c("x", "y") 
lvls <- c("vst")
SSvstdf$transformation <- factor(SSvstdf$transformation, levels = lvls)
ggplot(SSvstdf, aes(x=x, y=y))+geom_hex(bins=80)+
  coord_fixed()
```
I'm not sure how to interpret this... I'm not too convinced is good enough, but this is only for preliminary visualization
We might try the other one (I tried already, it wasn't that different)

Let's calculate the distance matrix between samples and plot it to have an idea of the grouping:
```{r}
SSDist <- dist(t(assay(SSDEseqvst)))
```

With an heatmap for example:

```{r}
library("pheatmap")
library("RColorBrewer")
SSDistM <- as.matrix( SSDist )
rownames(SSDistM) <- paste(SScolData$Condition)
colnames(SSDistM) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(SSDist,
         clustering_distance_rows = SSDist,
         clustering_distance_cols = SSDist,
         col = colors)
```
We would need an indication of what they are ... like names, but it's easier just look at the PCA:
```{r}
plotPCA(SSDEseqvst, intgroup=("Condition"))
```

Let's check more details on the samples:
```{r}
SSPCAdata <- plotPCA(SSDEseqvst, intgroup=c("Host", "Stage"), returnData=TRUE) ##ReturnData means do not plot but return them
SSpercentVar <- round(100*attr(SSPCAdata, "SSpercentVar")) ####Why is this empty?
ggplot(SSPCAdata, aes(PC1, PC2, color=Stage, shape=Host))+
  geom_point(size=3)+
  xlab(paste0("PC1: ", SSpercentVar[1], "% variance")) +
  ylab(paste0("PC2: ", SSpercentVar[2], "% variance")) +
  coord_fixed()
```

Let's make it nice for the poster:
```{r}
SSPCAdata <- plotPCA(SSDEseqvst, intgroup=c("Host", "Stage"), returnData=TRUE) ##ReturnData means do not plot but return them

SSntop=500
  SSv <- rowVars(assay(SSDEseqvst))
  SSselect <- order(SSv, decreasing = TRUE)[seq_len(min(SSntop,
                                                     length(SSv)))]
 ############## SSpca <- prcomp(t(assay(SSDEseqvst)[SSselect, ])) ###This is not necessary if I use SSPCAdata. Are these equivalent? I think so
  
  SSpercentVar <- SSPCAdata$sdev^2/sum(SSPCAdata$sdev^2)
  SSpercentVar <- round(SSpercentVar[1:2]*100,0)

  
SSpcaplot <- ggplot(SSPCAdata, aes(PC1, PC2, color=Stage, shape=Host))+
  geom_point(size=5)+
  xlab(paste0("PC1: ", SSpercentVar[1], "% variance")) +
  ylab(paste0("PC2: ", SSpercentVar[2], "% variance")) +
  scale_shape_manual(values = c("Copepod" = 15, 
                                "Free" = 17, 
                                "Stickleback" = 16,
                                "Bird" = 4)) +
  scale_color_manual(values = c("Growing" = "gray0",
                                "Infecting"= "firebrick",
                                "Reproducing"= "goldenrod3",
                                "Transmission" = "darkgreen",
                                "Growing_Transmission" = "cyan4")) +
  coord_fixed()+
  theme_bw()+
  theme(text = element_text(size = 15))+
  theme(panel.grid = element_blank(), panel.border = element_blank(), axis.line=element_line(), panel.background = element_rect(fill="transparent"), plot.background = element_rect(fill = "transparent"), legend.background = element_rect("transparent"), legend.key = element_rect(fill="transparent"))

############To save the pics:
#tiff(filename = "SSpca.tiff", width = 5, height = 4, units = "in", res = 1200, compression ="lzw")+
#png(filename = "SSpca.png", width = 5, height = 4, units = "in", res = 1200)  

#Or
#ggsave(SSpcaplot, 
#       filename = "SSpca.png",
#       bg = "transparent")
#But this doesn't allow to customize the resolution  
```
Why is it going the NA ?

Multi-dimensional-scaling MDS:
```{r}
SSmds <-as.data.frame(colData(SSDEseqvst))  %>%
         cbind(cmdscale(SSDistM))
ggplot(SSmds, aes(x = `1`, y = `2`, color = Stage, shape = Host, size=Pop)) +
  geom_point() + coord_fixed() + ggtitle("MDS with vst data")

```

To visualize some heatmap as additional explorative analysis:
```{r}
library("genefilter")
SStopvargenes <- head(order(rowVars(assay(SSDEseqvst)), decreasing = TRUE), 10) ##Select the 10 genes that bring more variance
SSmat <- assay(SSDEseqvst)[SStopvargenes, ]
SSmat <- SSmat - rowMeans(SSmat)
SS20genes <- as.data.frame(colData(SSDEseqvst)[,c("Stage", "Host")]) ##I can play with this filtering out the other pop and the samples from Hebert to see how it changes
pheatmap(SSmat, annotation_col = SS20genes)
```


########################################################END OF VISUALIZATION/QUALITY CHECKS ################################################

